---
title: "Exploratory Data Analysis"
output: html_document
date: "2022-11-29"
---

## Necessary libraries 

```{r}
#| warning: false
library(tidyverse)
library(dplyr)
library(readxl)
library(lubridate)
library(tidyr)
library(zoo)
library(janitor)
library(reactable)
library(data.table)
library(ggplot2)
```



## Importing data

We are importing construction cost index, Turnover rate, usd to tl exchange rate, House sales to foreigners, House Sales by Province data sets.
```{r}
ConsIndex <- readRDS("Datards/ConsIndex.rds")
Turnover <- readRDS("Datards/Turnover.rds")
House_sales_to_foreigners_month <- readRDS("Datards/House_sales_to_foreigners_month.rds")
House_sales_to_foreigners_year <- readRDS("Datards/House_sales_to_foreigners_year.rds")

monthly_usd_to_try <- read_excel("data/monthly_usd_to_try.xls")
monthly_usd_to_try$month <- as.Date(monthly_usd_to_try$month)
monthly_usd_to_try$month <- format(monthly_usd_to_try$month,"%Y-%b-%d")

PATH <- "data/House Sales by Provinces.xls"
house_sales_by_provinces <- read_excel(PATH, range = cell_rows(14:129), col_names = names(read_excel(PATH, skip = 2))) 

total_sales_by_year <- aggregate(house_sales_by_provinces[3], house_sales_by_provinces[1], FUN=sum)

House_sales_to_foreigners_month <- read_excel("data/House sales to foreigners.xls", 
                                                   sheet = "Month")
House_sales_to_foreigners_month$Month <- match(House_sales_to_foreigners_month$Month, month.name)

House_sales_to_foreigners_month$Date <- as.yearmon(paste(House_sales_to_foreigners_month$Year, House_sales_to_foreigners_month$Month), "%Y %m")

```
## Key Takeaways
1. When we examine the sales data, although the rate of sales to foreigners has increased over the years, it has remained around 5 percent. Considering this information, it seems difficult for the sale to foreigners to have a significant impact on the market.
2. After 2019, we can see that the annual average cost index exceeds the annual average turnover index.
3. When we examine the relationship between the dollar and TL exchange rate, we conclude that the TL has experienced a high depreciation against the dollar. And we observe that there is a clear correlation between the cost index and the dollar-tl exchange rate relationship.



### Total sales with Sales To Foreigners and Sales To Foreigners Share (%)
```{r}
ggplot(House_sales_to_foreigners_month, aes(Date)) + 
  geom_line(aes(y = Total, colour = "Total")) + 
  geom_line(aes(y = `Sales to foreigners`, colour = "Sales To Foreigners"))

ggplot(House_sales_to_foreigners_month, aes(Date)) + 
  geom_line(aes(y = `Share (%)`, colour = "Share (%)"))

```

Finding construction cost index yearly mean to compare

```{r}
ConsIndexYear <- ConsIndex %>% group_by(Year) %>% summarise(Construction_cost_mean = mean(index, na.rm = TRUE))
TurnoverYear <- Turnover %>% group_by(Year) %>% summarise(Turnover_mean = mean(`seasonal and calendar adjusted
Index`))

ConsIndexYear %>% left_join(TurnoverYear, by = "Year") %>% ggplot(aes(Year)) +
  geom_line(aes(y = Construction_cost_mean, colour = "Construction cost mean")) + 
  geom_line(aes(y = Turnover_mean, colour = "Turnover mean"))+
  labs(title="Construction cost and Turnover index yearly mean comparison")+
  xlab("Year") +
  ylab("Index")


```


```{r}
house_sales_by_provinces  <- house_sales_by_provinces  %>%
    fill(colnames(house_sales_by_provinces[1]), .direction = "downup")

# Creating Time Series Object for House Sales by Proviences
total_hs_ts <- ts(house_sales_by_provinces[3],start=c(2013,1),frequency=12)

options(repr.plot.width=21, repr.plot.height=8)
plot(total_hs_ts)
```

### Monthly usd to try graph and  Monhtly usd rate preprocessing

```{r}

monthly_usd_to_try <- read_excel("data/monthly_usd_to_try.xls")

monthly_usd_to_try$month <- as.Date(monthly_usd_to_try$month)
colnames(monthly_usd_to_try) <- c("date","value")

ggplot(data=monthly_usd_to_try, aes(x=date, y=value, group=1)) +
  geom_line()+
  geom_point()
```


### Monthly cost index and  Monhtly cost index preprocessing

```{r}

construction_cost_index <- read_excel("data/construction_cost_index_by_industries_and_cost_groups_preprocessed.xls")


construction_cost_index$date <- as.Date(construction_cost_index$date)
colnames(construction_cost_index) <- c("date","cost_type","value","material","labour")

cost_index = construction_cost_index %>% 
  filter(cost_type=="construction") %>%
  group_by(date) %>%
  summarize(value=sum(value))

ggplot(data=cost_index, aes(x=date, y=value, group=1)) +
  geom_line()+
  geom_point()
```




```{r}
# Yillik satista ulkeler
library(readxl)
HSTFBN <- read_excel("data/HSTFBN.xls", sheet = "FOREIGNERS BY NATIONALITIES")

total_sale_by_country <- HSTFBN %>% group_by(year,country) %>% summarise(sum=sum(total))


# 2017'den sonra Iran, Irak'i geciyor.
# 2021'den 2022'ye gecerken rusya da yuksek bir artis var.

ggplot(data=total_sale_by_country,aes(x=year,y=sum,fill=country)) +
  geom_bar(stat="identity",position = "dodge")
```




```{r}
house_sales_by_districts_preprocessed <- read_excel("data/house_sales_by_districts_preprocessed.xls")

house_sales <- house_sales_by_districts_preprocessed %>%
  group_by(year) %>%
  summarise(morgaged=sum(mortgaged), total=sum(total), first_hand=sum(first_hand), second_hand=sum(second_hand))


ggplot(data=house_sales,aes(x=year),group = 1 ) +
  geom_line(aes(y=morgaged,color='morgaged')) +
  geom_line(aes(y=total,color='total'))+
  xlab("Year")+
  ylab("Values")
```




```{r}
house_sales_by_districts_preprocessed <- read_excel("data/house_sales_by_districts_preprocessed.xls")

house_sales <- house_sales_by_districts_preprocessed %>%
  group_by(year) %>%
  summarise(morgaged=sum(mortgaged), total=sum(total), first_hand=sum(first_hand), second_hand=sum(second_hand))


ggplot(data=house_sales,aes(x=year),group = 1 ) +
  geom_line(aes(y=first_hand,color='first_hand')) +
  geom_line(aes(y=second_hand,color='second_hand')) +
  xlab("Year")+
  ylab("Values")
```




